### 2014-11-08 implementation.tex
rewrote threads/queue section.

Each node is protected by a main spinlock that allows threads to change node
attributes: incoming facts, owner thread, state variable and locality information.
A database spinlock is locked whenever the node enters into
the \textbf{working} state. If a node sends facts to another node placed in
another thread, the current thread first attempts to lock the database lock in order to update the
indexing structures of the target node, otherwise it adds the facts to the list of
incoming facts that are later processed by the owner thread.

@@

Threads pop active nodes from a thread local queue and execute new candidate rules to
generate new facts, which in turn can activate other nodes.
XXX NEEDS REWRITE
In order to reduce inter-thread communication, node priorities are implemented
at the thread level. This means that, when a thread picks the highest
priority node, it is actually the highest priority from the set of nodes owned
by the thread and not the highest priority node in the whole program.
To allow node priorities, we use 2 two pairs of queues per thread: a pair of doubly linked
lists known as the \emph{standard queue} and a pair of min/max heap known as the \emph{priority queue}.

The regular queue contains nodes without priorities and
supports push into tail, remove node from the head, remove arbitrary node, and
remove first half of nodes.
The priority queue contains nodes with priorities and is implemented as a binary
heap array. It supports the following operations: push into the heap, remove the min node,
remove an arbitrary node, remove half of the nodes (horizontal split), and
priority update.
Operations for removing half of the queue are implemented in order to support
node stealing, while operations to remove arbitrary nodes or update the
priority of nodes allows threads to change the priority of nodes.

@@

Both the regular and priority queue are actually implemented as 2 queues each.
The \emph{static queue} contains nodes that are local to a thread (and thus cannot be stolen) and
the \emph{dynamic queue} maybe accessed by other threads during node stealing.

>>>>

Each node is protected by a main spinlock that allows threads to
change node attributes: incoming facts, owner thread, state variable
and locality information.  A database spinlock is locked whenever the
node enters into the \textbf{working} state.

To avoid unneccesary copying, when a node sends facts to a node
located in another thread, the current thread first attempts to lock
the database lock of the target node in order to directly update its
indexing structures, otherwise it adds the facts to the list of
incoming facts that are later processed by the owner thread of the
target node.

@@

The main loop for each thread is to get an active node from one of its
local queues and process all applicable rules for that node generating
new facts which can result in new facts for the current node or other
nodes.  The thread will continue processing facts for the current node
until there are no new facts generated for it.  It will then pick
another active node.  If its queues are all empty it will try to steal
work (\ie nodes) from another thread.

To support priorities each thread has two pairs of queues: a pair of
doubly linked lists known as the \emph{standard queue} and a pair of
min/max heap known as the \emph{priority queue}.  The standard queue
contains nodes without priorities and supports push into tail, remove
node from the head, remove arbitrary node, and remove first half of
nodes.  The priority queue contains nodes with priorities and is
implemented as a binary heap array. It supports the following
operations: push into the heap, remove the min node, remove an
arbitrary node, remove half of the nodes (horizontal split), and
priority update.  Operations for removing half of the queue are
implemented in order to support node stealing, while operations to
remove arbitrary nodes or update the priority of nodes allows threads
to change the priority of nodes.

To minimize inter-thread communication, node priorities are
implemented at the thread level. Thus, when a thread picks the highest
priority node from the priority queue, it is only the highest priority
with respect to the set of nodes owned by the thread and not the
highest priority node in the whole program.

@@

Both the regular and priority queue are implemented as a pair of
queues.  This first of the pair is the \emph{static queue} which
contains nodes that cannot be stolen.  The other queue in each pair is
the \emph{dynamic queue} which contains nodes which can be stolen by
other threads.
